Study of the hadronization of charged pions

### 2019-01-23 1-Answers-Review.tex

\section{Larger Issues }
1.
Ownership of the pi+ channel.  This was brought up in an email from the committee chair 
(M. Wood).  Before this analysis can move on to round 2 of the review, the ownership issue 
has to be resolved.  The NPWG chair (also M. Wood) has left it up to the eg2 group to 
resolve it.

{\it We are aware of this issue and we started regular meetings between the different 
groups toward the objective of publishing as soon as possible results coherent with
the different analysis performed on this topic (all three pions in particular). This
will eventually translate into an addendum to this note with comparisons between Hayk's 
and this analysis, which will be used to confirm the proper sizing of the systematic errors. 
In a later stage an extra analysis note from Hayk will be submitted
for the extraction of multi-dimensional results on $\pi^+$.} \\

2.
The particle ID and cuts seem to be well understood.  The acceptance correction and 
radiative corrections need to be developed. There is confusion about how the acceptance 
corrections were done (see Section 2 below). The radiative corrections are incomplete as 
reported in the note.  Both corrections need to be well-understood before the committee can 
approve of the analysis.

{\it The text about acceptance has been reviewed in details to be clearer in this new 
version. The section about radiative corrections has been updated with the new method.
We can have a meeting to clarify some of the simpler issues more rapidly.} \\

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=7.5cm] {answer-fig/VertexC.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexAl.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexFe.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexSn.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexPb.png} 
\caption {Vertex origin along $z$ (in cm) of electrons as a function of run 
number for the different solid targets.}
\label{VertexSolid}
\end{figure}

\begin{figure}[tbp]
\centering
\includegraphics[width=7.5cm] {answer-fig/VertexDeutC.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexDeutFe.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexDeutSn.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexDeutPb.png} 
\caption {Vertex origin along $z$ (in cm) of electrons as a function of run 
number for the deuterium target in different solid targets configurations.}
\label{VertexLiquid}
\end{figure}

6.
P17, Fig 12 - what solid target is shown?  Are the shifts the same for electrons and pions?

{\it Sadly, this cannot be checked, the kumac used for this figure specify a file name that
do not exist anymore (codes and files have been moved at some point and non essential files
have been removed in the process). It is always possible to reproduce the figure with a root 
routine if the reviewer feel it is necessary to know the specific run used.

@@

7.
P18, Start of section 2.2.5 - I’m not sure I understand the logic in this.  If you have a drift 
chamber problem, it will be manifest in both the solid and liquid target data and the ratio of 
these should be relatively stable.  Why do you not determine run quality by looking at the 
event rate per coulomb of beam current?  For this quantity, a change in the detector 
operation will show up.

{\it You are right, the motivation for this cut is not stated properly. We are looking 
at the ratio because it would affect our result, while ignoring issues that would not
affect our observable. It is important to note that the initial run list was already
cleaned from known detector failures. (The text is corrected in the note.) }

8.
P19, FIg. 14- Hard to tell how much data was discarded. Maybe give a table of means and 
sigmas for each target or show the cuts on the figure are horizontal lines.

{\it The quantity of data removed is difficult to assess, since the problematic runs were
not passed through the final analysis. We were unable to recover the orginal values of the
fits for these specific figures and we 
lack the problematic run which have been deleted. At this stage, it would be rather time
consuming to recover the raw data and update all the softwares to run on the new machines, so 
we did not produce this figure. If the reviewer think it is necessary to update this 
figure and provide more information about the discarded data let us know, it is time 
consuming but of course possible.}

@@

11.
P19, eq. 7 - Is it really just the sum of weights or are they the counts multiplied by the 
weights?

{\it They are indeed the counts multiplied by the weights.
The sum indicated runs over all the recorded events, it is not associated to the weight
bins or anything like that. We adapted the language in the note to make it clearer.} \\

12.
P23, I’d be very worried about the MC modeling acceptance on falling edges like that at the 
right of the top right plot.

{\it It is indeed a problematic issue, linked to the abscence of radiative effects in the 
simulation. At the time that it was generated, no better simulation was available. This
discrepency in $\nu$ motivates to also use it for the binning of the acceptance. This way,
it limits the impact of this difference.} \\

@@

13.
P25, I find this discussion quite confusing.   At top you talk about two binnings.  It’s not just 
binning that is changing, right?  You are using different variables.  Then you talk about the 
extracted weight, but that weight is defined with only one of those sets of variables.  You 
also do it for the other set?  Then you an arbitrary weight cut to take care of the problems 
listed in the middle of the page?  Why not just increase the size of the bins in regions where 
you have too few events or large bin migration?  I don’t see how the procedure you use 
does not lead to holes in your data somehow.  And I don’t really understand what you are 
doing with the second weighting factor.  How is this not doing a double correction for 
acceptance, i.e., a double counting?  This whole section needs to be reworked.  I am fairly 
certain that I don’t really understand what you are doing but it leaves me worried that 
something is amiss.  Anyway, that’s not where you want readers to be.

{ \it The section has been reworked to clarify what we are doing. Few points you raised 
are also answered here:

The first set of variables is the one used for the acceptance correction presented in the results,
while the second set is only used to evaluate the systematic errors associated with the correction. In the 
note, the formulas are generally given only for the first set for simplicity.

The motivation for using such a strange cut based on the product of different unrelated
values is actually directly linked to the problem of inbent 
particles being directed towards the beam pipe. The $\pi^-$ acceptance drops very 
strongly at low $p_\perp^2$ leading to very high weights, however there is no problem of 
bin migration
or low statistics because the generated yield is very high in this region. We found that 
the best way
to keep these bins while discarding bins where the size of the weight is less problematic 
but with issues concerning bin migration and statistics was to apply this type of cut.

In term of method, the two
step acceptance correction is hopefully going to be clearer. It is the same as the method used 
in the previous L. El Fassi et al. analysis note of eg2 data. The second step of the procedure is
specifically used to take care of the bins that are removed by the cut on weights. You 
can see in the formula that the weights
of the previously calculated acceptance are used in this second step, ensuring that we do not
double correct the data.

@@

The size of the bins is not constant and is indeed increased in region with small event yield.
However, the multi dimensional binning cannot be fully adapted in a general maner. On this
topic we added tables in the note indicating the bin limits that we used.} \\

14.
P26, bottom, if you have regions with very low acceptance, should they be cut out?

{\it In the case of the $\Delta P_\perp^2$, we need to measure an average of the
transverse momentum distribution. Completely cutting a part of the distribution
would indeed solve the acceptance problems for $\pi^-$ but would make the $\langle p_\perp^2 \rangle$
measurement meaningless. We therefore have to live with it and do our best to
reconstruct the whole distribution even if it means having larger error bars.} \\

@@

15.
P26 - "consistent with HERMES results" - could you add HERMES points to the plot with 
your data points to show what you mean as consistent?

{\it This sentence states that similarly to HERMES, we find that both charged pions have similar
transverse momentum. This can be seen in the Figure 4 of the note where we show the result from 
HERMES. We did not want to mislead the reader into any claim about compatibility between 
HERMES and CLAS data here, so we reworded the sentence. We want to point out that such claim 
would be highly model dependent since it is expected that the effect varies with energy. } \\

16.
P27 fig.22-
 It's unclear to which point the error bar is associated. I think the common 
practice is to slightly shift markers horizontally.

{\it Sorry, we overlooked this for these intermediary plots.} \\

@@

17.
P27, Fig. 22 - the right column of plots is not the same as those in Fig. 15.  Why change the 
variables?

{\it Sorry, it is a mistake on our part, there is no reason.} \\

@@

18.
P28 - What is Coulolmb Correction applied to? Momentum of charged particles? Could you 
plot effect of this correction on particles? Is it really big enough to be visible within our 
resolutions ?

{\it The Coulomb correction is applied to incoming and outgoing electrons as well as
produced charged pions. Indeed the effect is small in most of the phase space, but 
by slightly shifting the $z$ distribution it slightly changes the normalization. A
point to point effect is also visible for low $z$ pions. Note that in general, there
is also a focussing factor in Coulomb corrections, but they happen to cancel out in
our specific observables.} \\

@@

19.
P30 - 
fig.24 is missing X axis label. You also may consider 2D histogram for better display.

{\it This figure is replaced by others in the new radiative correction section.} \\

20.
P31 - I don’t understand the isospin correction.  It does smooth out the data in Fig. 27; 
however, I am missing the physics motivation.  Why does it matter if the quark came from a 
proton or a neutron?  I think you just clarify the explanation.

{\it This correction is here to take care of the u/d quark imbalance in heavier nuclei.
Indeed the multiplicity of positive pions on neutrons is significantly smaller than on protons and that
should not be mistaken with absorbtion in the nuclear medium. We clarified the text in the
note.} \\

@@

21.
P33, Sec. 2.5.1 - list the negligible effects.

{\it The sentence has been revised to list all effects tested.} \\

@@

22.
P34, Fig. 28 - why is the resolution poorer than in Fig. 12?

{\it Resolution is the same, it looks wider because, on this figure, we show the integrated 
result over the 6 sectors, which are slightly shifted between each other.} \\

@@

23.
P34, why do you have a 1\% normalization error in all the multiplicity ratios when you say just 
above that for this particular sample with two particles in the final state the events in the test 
regions dropped to 0.01?  Doesn’t that mean the target contamination normalization error is 
smaller than 1\%?

{\it Multiplicity ratios are a double ratio. The claim is that the semi-inclusive part 
is not affected, but the inclusive electron ratio is affected. This is included
as a normalisation error. We revisited the paragraph to make this point clearer.} \\

@@

24.
P34, with this technique of comparing results using the two different sets of variables (you 
call it two binnings, but I think that is sort of misleading), you are using the same events.  
Does this not mean you have a statistical contribution to this systematic error?  You say you 
noticed larger systematic errors for the pi- events due to lower acceptance and bigger 
weights.  Might it also be lower statistics and so you have statistical fluctuations contributing 
to the systematic error?  If so, you are sort of double-counting the statistical error and 
inflating the systematic error more than it should be.

{\it This is an interesting question and we might have overlooked this problem. Looking into the
issue we noticed an error in our function calculating the statistical errors. This change 
would modify our initial statement. Altogether, after reviewing the question, we believe that 
because the different acceptance are calculated from the same simulated data and applied to 
the same experimental data, our systematic error calculation should not include the effect 
of statistical errors. In conclusion, we removed the previous statement and now fully include 
the systematic errors for the transverse momentum broadening.} \\

@@

25.
P35 - It would clarify the discussion with a summary or table of the point-to-point systematic 
errors.

{\it The point-to-point errors include only contributions from the particle 
identifications and acceptance effects. It is also different for every bin, so
we did not think a table would be useful to summarize only two contributions.} \\

26.
P35 -
 add plots with systematic uncertainty bands on it to demonstrate bin-by-bin values.

{\it We add this on the figures intended to be published.} \\

@@

27.
P35, end of top paragraph, you say you don’t include the acceptance systematic, but it looks 
to me like it is included in Table 8?  What am I missing?

{\it We decided to keep the normalization error, not the point to point. This was not
said clearly in the text, it is now changed as mentioned above.} \\

@@

28.
P36, top paragraph, “that a factorization holds” means what??  Should explain.

{\it This sentence has been reformulated with an additional reference.} \\

29.
P36 top of section 3.1.1, 5\% versus 5\% looks more like 1 sigma rather than the 1.5 sigma 
you say in the text.

{\it The reson we quote 1.5, is that one need to move $\pi^+$ one sigma down and $\pi^-$ one
sigma up. So for the overall normal deviation (sigma) we quadratically sum them,
leading to 1.5, you think it should say 2?} \\

### 2019-01-23 chap5.tex
Correct the Delta beta formula
Finalized the answers

The previous analysis from L. El Fassi et al. used $\pm 0.05$ ($\sim 3\sigma$), we decided
to reduce this number to about $2\sigma$ because of the difference of observables. The 
color transparency analysis was reconstructing the $\rho^0$ and implemented strong kinematical
cut that reduced significantly possible contaminations. Our measurement uses all the 
semi-inclusive production and is therefore more sensitive to contaminations. See next 
question for profiles.

@@

\begin{subequations}\label{TOF-Pip}
\begin{align}
    p_\pi &\leq 3.375 \text{ GeV/c : }
  & \Delta \beta &> \max\left(-0.03,{p \over\sqrt{p^2+0.4^2}}\right), \\ 
    p_\pi &> 3.375    \text{ GeV/c : }
  & \Delta \beta &> \max\left(-0.02,{p \over\sqrt{p^2+0.7^2}}\right).
\end{align}
\end{subequations}

The much more apparent contamination sources than for negative pions, allow to study 
the TOF profile much more thoroughly. You can see in figures~\ref{TOF-1}, \ref{TOF-2} and 
\ref{TOF-3} the $\Delta \beta$ profiles for many momentum slices. In figure~\ref{TOF-1}
one can see some contamination from lighter leptons ($e+$ and $\mu^+$). These represent
a very small amount and one can see some justification for the choice of a $\pm 0.03$
cut as these particles do not appear to be discernable beyond that level. Similarly
in figure~\ref{TOF-2} one can see the contribution from kaons. The larger contribution
of this contamination motivated the tightening of the TOF cut on this side, to the limit
where kaons are a minority in the tail of the pion distribution. Finally a similar 
method is applied for the even larger proton contamination as shown in figure \ref{TOF-3}.

@@

@@

After testing several different acceptance correction methods, we finally 
converged with the one using four dimensions and a limited number of bins.
The two different approaches to binning the data presented on a table 
\ref{tab:AcceptBinning} are used to evaluate the systematic errors. The 
difference between these two methods on iron data are used to evaluate 
the relative uncertainties presented in a table~\ref{tab:SysAcc}. We noticed 
the significantly large errors for $\pi^-$s, which were expected, due to their 
lower acceptance and larger weights.

Systematic errors are also
considerably larger for \dpt measurements. This is mainly due to the nature of 
the \dpt observable; as it is a difference, it enhances relative errors 
significantly. Also, as shown in figure~\ref{fig:AcceptPlots}, error bars 
became much larger for the corrected \dpt from the acceptance effect. This 
result indicated an important statistical sensitivity introduced by our 
weighting methods. The errors on \dpt are of the same order, or smaller, than 
the differences observed between the two weighted data-sets. This is an 
indication that the uncertainties reported in table \ref{tab:SysAcc} are, in 
nature, similar to the statistical errors, in the sense that both originate 
from the large variation between event-by-event weights. For this reason, we 
decided not to include this error in our systematic budget to avoid double 
counting this effect.

>>>>

The previous analysis from L. El Fassi et al. used $\pm 0.05$ ($\sim 3\sigma$), we decided
to reduce this number to about $2\sigma$ because of the difference of observables. The 
color transparency analysis was reconstructing the $\rho^0$ and implemented strong kinematical
cut that reduced significantly possible contaminations. Our measurement uses all the 
semi-inclusive production and is therefore more sensitive to contaminations.

@@

\begin{subequations}\label{TOF-Pip}
\begin{align}
    p_\pi &\leq 1.838 \text{ GeV/c : }
  & \Delta \beta &> \max\left(-0.03,{p \over\sqrt{p^2+0.4^2}}- {p \over \sqrt{p^2 + m_\pi^2}}\right), \\ 
    p_\pi &> 1.838    \text{ GeV/c : }
  & \Delta \beta &> \max\left(-0.02,{p \over\sqrt{p^2+0.7^2}}- {p \over \sqrt{p^2 + m_\pi^2}}\right).
\end{align}
\end{subequations}

The much more apparent contamination sources than for negative pions, motivates us to study 
the TOF profile much more thoroughly. One can see in figures~\ref{TOF-1}, \ref{TOF-2} and 
\ref{TOF-3} the $\Delta \beta$ profiles for many momentum slices. In figure~\ref{TOF-1}
one can see some contamination from lighter leptons ($e+$ and $\mu^+$). These represent
a very small amount and one can see some justification for the choice of a $\pm 0.03$
cut as these particles do not appear to be discernable beyond that level. Similarly
in figure~\ref{TOF-2} one can see the contribution from kaons. The larger contribution
of this contamination motivated the tightening of the TOF cut on this side, to the limit
where kaons are a minority in the tail of the pion distribution. Finally a similar 
method is applied for the even larger proton contamination as shown in figure \ref{TOF-3}.

@@

\subsubsection{Endcap correction}

TO BE IMPLEMENTED IN THE ANALYSIS

The goal of this work is to correct the measured multiplicity ratio:
\begin{equation}
  R_h^A |_{corr} = {N_h^A / N_e^A \over (N_h^D-N_h^{EC}) / (N_e^D-N_e^{EC})},
\end{equation}
with the $EC$ target being the endcaps of the liquid target. Using a Taylor expension at first order for small $N^{EC}$, we find that:
\begin{equation}
  {N_h^D-N_h^{EC} \over N_e^D-N_e^{EC}} = {N_h^D \over N_e^D} - {N_h^{EC} \over N_e^D} + {N_e^{EC} N_h^D \over {N_e^D}^2}.
\end{equation}
After few straightforward transformation, we get
\begin{equation}
  R_h^A |_{corr} = R_h^A \left( 1 - {N_e^{EC} \over N_e^D} \left( 1-R_h^{Al} \right) \right).
\end{equation}
Based on target thickness ${N_e^{EC} \over N_e^D} \sim 0.03$ and $R_h^{Al} \sim 0.65$, the correction is on average a 1\% decrease of the measured multiplicity ratio.

With the same method, the transverse momentum broadening can be corrected with:
\begin{equation}
  \Delta P_T^2 |_{corr} = \Delta P_T^2 + R_h^{Al} {N_e^{EC} \over N_e^D} \Delta P_T^2|_{Al}
\end{equation}

@@

After testing several different acceptance correction methods, we finally 
converged with the one using four dimensions and a limited number of bins.
The two different approaches to binning the data presented on a table 
\ref{tab:AcceptBinning} are used to evaluate the systematic errors. The 
difference between these two methods on iron data are used to evaluate 
the relative uncertainties presented in a table~\ref{tab:SysAcc}. We noticed 
the significantly large errors for $\pi^-$s, which were expected, due to their 
larger acceptance weights. Systematic errors are also
considerably larger for \dpt measurements. This is mainly due to the nature of 
the \dpt observable; as it is a difference, it enhances relative errors 
rather than cancelling them.
